Fusion Moves for Correlation Clustering

### 2014-11-05 paper/egpaper_for_review.tex
work on introduction

\listoftodos[Notes]
%%%%%%%%% ABSTRACT
\begin{abstract}
   We address the problem of partitioning a  graph
   into a previously unknown number of clusters.
   Among all partitions, the one with the minimal 
   sum of of cut edge weights is chosen. 
   This problem is known as correlation clustering 
   or multicut.
   We propose a general framework to find
   high quality approximate solutions for 
   this NP-hard problem based on a move making algorithm:
   We use candidate solutions from a proposal generator
   to iteratively improve the best observed solution similar
   to fusion moves.
   The proposed solver outperforms any other solver
   w.r.t. to any time performance.
   The solutions found by this solver are close
   to global optimal solutions w.r.t. energy
   and problem specific measurements as VI for
   image segmentation problems.

@@

\paragraph{Our Notation}
$G$ Graph.
$V$ Set of Nodes.
$E$ Set of Edges.
$l$ node label.
$y$ edge label.
$w$ edge weight
$MC(G)$ multicut polytope
$P(G)$ the set of valid edge labeling (MC(G) is convex hull of P(G))
$y^*$ for opt. labeling
$G_y=(V_y,E_y)$ contradiction of graph $G$ with respect to $y$

%$\rho_y : E \to E_y$
%$\rho_y^{-1} : E_y \to E$

$\rho_y : P(G) \to P(G_y)$

$\rho_y^{-1} : P(G_y) \to P(G)$
\begin{itemize}
  \item Multicut in computer vision, applications, sparse segmentation -> IMPORTANT PROBLEM
  \item Problem: Existing Methods does not scale, even CGC. - Large scale 3d problems
  \item Contribution: Fast scalable method using novel fusion moves for correlation clustering
\end{itemize}

Given an edge weighted graph, positive weights (\emph{attractive})
encourage the adjacent nodes to be in the same 
connected component, while negative weights (\emph{repulsive}) encourage
the nodes to stay in different connect components.
Therefore the sign of the weights encodes if two
nodes should be merged or not and the magnitude of the weights encodes
the certainty of this desire.
The objective of correlation clustering / multicuts
is finding the cut with a minimal sum of cut edge weights.
The number of connected components / clusters is discovered
from the weights ( rewrite?!?).
Correlation clustering / the multicut is NP-hard \cite{???}.
%
Despite the NP-hardness, correlation clustering has been 
successfully used for
\begin{inparaenum}[(i)]
    \item partitioning a superpixel region adjacency graph~\cite{andres_2011_iccv,kroeger_2012_eccv}
    \item with optional long range repulsive edges~\cite{andres_2013_emmcvpr}.
    \item Alush and Globerger showed how to average multiple segmentations with the multicut objective~\cite{alush_2012_pami}.
    \item Multicuts can also be used for interactive segmentation~\cite{bagon_2011_arxiv},
    \item for co-segmentation~\cite{glassner_2011_cvpr}
    \item and to cluster sparse and graphs~\cite{???}.
\end{inparaenum}

Despite the nice properties of correlation clustering
to perform a \say{one-shot agglomeration of supervoxels} ~\cite{nunez_iglesias_2013},
the main draw back of correlation clustering is scalability~\cite{nunez_iglesias_2013}.
Existing solvers for correlation clustering do not scale for non planar graphs.
Even state of the art approximate move making algorithm fail 
to give good approximations non planar graphs.
When increasing the problem sizes, either the runtime explodes or the 
approximations are far away from being useful.

\textbf{Contribution:}
\begin{inparaenum}[(i)]
\item Within this work we propose a fast and scalable move making algorithm for correlation clustering,
\item which generalized  fusion moves \cite{???} to correlation clustering problems.
\item We reduce the inference to a series of moves where
each moves optimizes over a subspace spanned by the current best solution
and a proposal solution.
\item We propose two versatile proposal generators,
\item and show how to optimize these moves.
\item We give a polyhedral interpretation of this algorithm,
\item and evaluate the proposed method 
on existing benchmark problems.
\item We show state of the art any time performance on those instances.
\end{inparaenum}[(i)]

@@

\section{Problem Formulation}\label{sec:problem_formulation}
Def: Graph, weighted graph, cut, multicut

\begin{center}
    \begin{eqnarray}
        l^* &=& \argmin_{L \in \{1,\ldots,|V|\}^{|V|}} \sum_{ (i,j) \in E } w_{ij} \cdot [l_{i} \neq l_{j}] \label{eq:nodeproblem}\\
   %     y_{ij}^* &=& [l_{u} \neq l_{v}] 
        y^* &=& \argmin_{y \in MC(G)\cap \{0,1\}^{|E|}} \sum_{ (i,j) \in E } w_{ij} \cdot y_{ij} \label{eq:edgeproblem}%\\ 
    \end{eqnarray}
\end{center}

The exist an surjective mapping from a node-label $l$ to edge-labeling $y$ and
a bijective mapping from a partitioning of $V$ to vertices of the multicut polytope $MC$.
Consequently, 
(i) problems \ref{eq:nodeproblem} and \ref{eq:edgeproblem} are equivalent
and (ii) the node-labeling is not unique for a given partitioning and introduce some
ambiguities.

>>>>

%%%%%%%%% ABSTRACT
\begin{abstract}
   We address the problem of partitioning a  graph
   into a previously unknown number of clusters.
   Among all partitions, the one with the minimal 
   sum of of cut edge weights is chosen. 
   This problem is known as correlation clustering 
   or multicut.
   We propose a general framework to find
   high quality approximate solutions for 
   this NP-hard problem based on a move making algorithm:
   We use candidate solutions from a proposal generator
   to iteratively improve the best observed solution similar
   to fusion moves.
   The proposed solver outperforms any other solver
   w.r.t. to any time performance.
   The solutions found by this solver are close
   to global optimal solutions w.r.t. energy
   and problem specific measurements as VI for
   image segmentation problems.

@@

Segmentation is an important problem in computer vision and data mining as a first step
towards understanding an image or data in general. 
Many algorithms start with an over-segmentation
into superpixels (superregions), which are then clustered into ``perceptually meaningful''
regions.
Usually, the number of these regions is \emph{not} known beforehand.

Recently, the multicut formulation~\cite{chopra_1993_mp, Kappes-2013-multicut}
(sometimes called \emph{correlation clustering}, \cite{bansal_2004_ml})
has become increasingly popular for unsupervised image segmentation.
%
Given an edge-weighted region adjacency graph,
the problem is to find the segmentation which
minimizes the cost of the cut edges.
Such an approach has been shown to yield
state-of-the-art results on the Berkeley Segmentation Database
\cite{andres_2011_iccv,yarkony_2012_eccv,alush_2013_simbad}.

Despite the clear mathematical formulation and nice properties of correlation clustering
to perform a one-shot agglomeration of supervoxels,
the main draw back of correlation clustering is scalability~\cite{nunez_iglesias_2013}
which arise from the NP-hardness of the problem.

Recently, Beier~\etal~\cite{Beier-2014} presented a promising 
approximative method called Cut, Clue \& Cut.
While this scales well for planar graphs, it does not for non planar graphs and has a bad any-time performance.
A further inherent drawback of this method is that the subproblems are not limited in their size.
%
Consequently, partition problems on large scale data, \eg
huge volume images in computational neuroscience~\cite{kroeger_2012_eccv}
or social networks~\cite{Leskovec-2010}, 
are not tractable, because the runtime explodes or the 
approximations are far away from being useful.

In the present work we present a novel approach that uses the concept of fusion for correlation clustering.
Starting with an initial segmentation we fuse the current segmentations with a proposed one.
Due to the ambiguity of node-labeling for unsupervised segmentations~\cite{} 
original fusion moves~\cite{Lempinski-2010} are improper, see Sec.~\ref{sec:fusion} for discussion.
To overcome this problem we suggest to use fusion moves on the edge-variables that indicates a cut and form a
bijective representation of segmentations.
The subproblem of fusing of segmentations reduce to a much smaller multicut problem which can be solved extremely efficiently.

\textbf{Contribution:}
\begin{inparaenum}[(i)]
\item We propose a fast and scalable move making algorithm for correlation clustering,
\item and give a polyhedral interpretation of this algorithm.
\item We propose two versatile proposal generators, 
\item and evaluate the proposed method on existing and new benchmark problems.
\end{inparaenum}

% Given an edge weighted graph, positive weights (\emph{attractive})
% encourage the adjacent nodes to be in the same 
% connected component, while negative weights (\emph{repulsive}) encourage
% the nodes to stay in different connect components.
% Therefore the sign of the weights encodes if two
% nodes should be merged or not and the magnitude of the weights encodes
% the certainty of this desire.
% The objective of correlation clustering / multicuts
% is finding the cut with a minimal sum of cut edge weights.
% The number of connected components / clusters is discovered
% from the weights ( rewrite?!?).
% Correlation clustering / the multicut is NP-hard \cite{???}.
% %
% Despite the NP-hardness, correlation clustering has been 
% successfully used for
% \begin{inparaenum}[(i)]
%     \item partitioning a superpixel region adjacency graph~\cite{andres_2011_iccv,kroeger_2012_eccv}
%     \item with optional long range repulsive edges~\cite{andres_2013_emmcvpr}.
%     \item Alush and Globerger showed how to average multiple segmentations with the multicut objective~\cite{alush_2012_pami}.
%     \item Multicuts can also be used for interactive segmentation~\cite{bagon_2011_arxiv},
%     \item for co-segmentation~\cite{glassner_2011_cvpr}
%     \item and to cluster sparse and graphs~\cite{???}.
% \end{inparaenum}

% Despite the nice properties of correlation clustering
% to perform a \say{one-shot agglomeration of supervoxels} ~\cite{nunez_iglesias_2013},
% the main draw back of correlation clustering is scalability~\cite{nunez_iglesias_2013}.
% Existing solvers for correlation clustering do not scale for non planar graphs.
% Even state of the art approximate move making algorithm fail 
% to give good approximations non planar graphs.
% When increasing the problem sizes, either the runtime explodes or the 
% approximations are far away from being useful.

% \textbf{Contribution:}
% \begin{inparaenum}[(i)]
% \item Within this work we propose a fast and scalable move making algorithm for correlation clustering,
% \item which generalized  fusion moves \cite{???} to correlation clustering problems.
% \item We reduce the inference to a series of moves where
% each moves optimizes over a subspace spanned by the current best solution
% and a proposal solution.
% \item We propose two versatile proposal generators,
% \item and show how to optimize these moves.
% \item We give a polyhedral interpretation of this algorithm,
% \item and evaluate the proposed method 
% on existing benchmark problems.
% \item We show state of the art any time performance on those instances.
% \end{inparaenum}[(i)]

@@

\section{Notation and Problem Formulation}\label{sec:problem_formulation}
Let $G=(V,E, w)$ be a weighted graph of nodes $V$ and edges $E$.
%
The function $w : E \rightarrow \mathbb{R}$ assigns a weight to each edge.
A positive weight expresses the desire that two adjacent nodes should
be merged, whereas a negative weight indicates
that these nodes should be separated into two different regions.
%
A \emph{subgraph} $G_A = \{A, E_A, w\}$ consists
of nodes $A \subseteq V$ and edges $E_A := E\cap (A\times A)$.
%
A segmentation of the graph $G$ can by either given by an 
node labeling $l \in \mathbb{N}^{|V|}$
or an edge labeling $y \in\{0,1\}^{|E|}$.  
An edge labeling is only consistent if it contains no dangling edges~\cite{Kappes-2013-multicut}.
We denote the set of all consistent edge labelings by $P(G)$.
The convex hull of this set is known as the \emph{multicut polytope} $MC(G)$.

A consistent edge labeling $y$ defines a segmentation of $V$ by enumerating the connected components 
in $G|_y=(V,\{e\in E|y_e=0\}$ by $l_v = \#(v)$, where $\#(v)$ denotes the number of the connected component of the node $v\in V$.

We define the contradiction of graph $G_y=(V_y,E_y)$
with $V_y=\{\#(v)|v\in V\}$ and $E_y=\{\#(v),\#(u)|uv\in E, y_{uv}=1\})$.

%$\rho_y : P(G) \to P(G_y)$
%$\rho_y^{-1} : P(G_y) \to P(G)$

% \begin{itemize}
%   \item Multicut in computer vision, applications, sparse segmentation -> IMPORTANT PROBLEM
%   \item Problem: Existing Methods does not scale, even CGC. - Large scale 3d problems
%   \item Contribution: Fast scalable method using novel fusion moves for correlation clustering
% \end{itemize}

% Def: Graph, weighted graph, cut, multicut

Given a weighted graph $G=(V,E,w)$ we consider the problem of segmenting $G$ such that the costs
of the edges between distinct segments is minimized. This can be formulated in the node~\ref{eq:nodeproblem} 
or edge domain~\ref{eq:edgeproblem}

\begin{eqnarray}
  l^* &=& \argmin_{L \in \mathbb{N}^{|V|}} \sum_{ (i,j) \in E } w_{ij} \cdot [l_{i} \neq l_{j}] \label{eq:nodeproblem}\\
  % y_{ij}^* &=& [l_{u} \neq l_{v}] 
  y^* &=& \argmin_{y \in P(G)} \sum_{ (i,j) \in E } w_{ij} \cdot y_{ij} \label{eq:edgeproblem}%\\ 
\end{eqnarray}

% The exist an surjective mapping from a node-label $l$ to edge-labeling $y$ and
% a bijective mapping from a partitioning of $V$ to vertices of the multicut polytope $MC$.
% Consequently, 
% (i) problems \ref{eq:nodeproblem} and \ref{eq:edgeproblem} are equivalent
% and (ii) the node-labeling is not unique for a given partitioning and introduce some
% ambiguities.
